---
title: "Findings for Koppers invesitgation with Cicero Independiente"
execute:
  output: false
  echo: false
format:
  html:
    code-tools: true
    theme: cosmo
    toc: true
---

```{r}
library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(stringr)
library(DT)
```

```{r}
# LOAD DATA 

# Pollutant codes for Illinois, which seems to have it's own code system. These were copy and pasted from the following website: https://epa.illinois.gov/topics/air-quality/planning-reporting/annual-emission-reports/reference-tables/pollutants.html
pollutant_code_crosswalk <- read_csv(here("data", "manual", "pollutant_code_crosswalk.csv")) %>% 
  clean_names()

# These are the total annual emissions reported to the Illinois EPA in the Air Emissions Invetory. We recieved this data through an open records request to the IEPA 
cook_county_emissions <- read_excel(here("data", "raw", "foia_reported_emissions_2012-2021.xlsx")) %>% 
  clean_names() %>% 
  left_join(pollutant_code_crosswalk, by = "pollutant_code") %>% 
  pivot_longer(cols = contains("x"), names_to = "year", values_to = "emissions") %>% 
  mutate(year = str_replace(year, "x", ""))

num_facilities <- cook_county_emissions %>% 
  distinct(name, address)

num_pollutants <- cook_county_emissions %>% 
  distinct(pollutant_code)

# National data from the EPA's Toxic Release Inventory on emissions for benzene, naphthalene, phthalic anhydride, maleic anhydride, quinoline, styrene and creosote. This data is the first file of the Basic Plus dataset. I pulled the years 2012 to 2022 since that is what we have for the Cook County emissions inventory. The data loaded here was processed in cleaned in the etl/prep_tri_data.R file. 
koppers_tri_pollutants <- read_csv(here("data", "processed", "koppers_tri_pollutants.csv"))

# This data comes from the EPA's AirToxScreen and represents the additional amount of people who could have cancer from this specific pollutan, e.g. an extra 10 in one million
cancer_risk_tract <- read_csv(here("data", "processed", "airtoxscreen_naphthalene_benzene_risk.csv")) %>% 
  clean_names() 
  
```


## How Koppers compares to other polluters in Cook County 

### Data
- These are the total annual emissions reported to the Illinois EPA in the Air Emissions Invetory. We received this data through an open records request to the IEPA 

### Findings 
- Koppers reports two dozen pollutants to the EPA, including several irritants and hazardous pollutants as well as five known or possible carcinogens 
- The pollutants below are a list of the top emissions of Koppers, exluding carbon and particulate matter categories, and including any known or possible carcinogen
- 0f 41 pollutants emitted in Cicero, Koppers is the highest polluter of 22 of them. The next top polluter is only at the top of the list for 7 pollutants
 

```{r}
cicero_mean <- cook_county_emissions %>% 
  filter(city == "Cicero") %>% 
  group_by(name, address, pollutant_code, chemical_name) %>% 
  summarize(mean_tons = mean(emissions, na.rm = TRUE), mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2))

```

```{r}
cicero_mean_highest_values <- cicero_mean %>% 
  group_by(pollutant_code, chemical_name) %>% 
  filter(mean_pounds == max(mean_pounds))
              
```


### Cicero's average emissions in the last decade by polluter and pollutant
```{r}
#| output: true
datatable(cicero_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```

### Cicero's top polluters in the last decade by polluter and pollutant
```{r}
#| output: true
datatable(cicero_mean_highest_values, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```



```{r}
koppers_pollutants <- cook_county_emissions %>% 
  filter(name == "Koppers Inc") %>% 
  group_by(name, address, pollutant_code, chemical_name) %>% 
  summarize(mean_tons = mean(emissions, na.rm = TRUE), mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) 

```

### Kopper's total emissions by pollutants from 2012 to 2021
```{r}
#| output: true
datatable(koppers_pollutants, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```




```{r}
cook_county_so2_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "SO2") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_nox_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "NOX") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_vom_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "VOM") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_panhy_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "85449") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_manhy_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "108316") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_naphthalene_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "91203") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_benzene_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "BENZENE") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000, 4)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```

```{r}
cook_county_quinoline_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "QUINOLINE") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```

```{r}
cook_county_styrene_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "STYRENE") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


```{r}
cook_county_creosote_avg <- cook_county_emissions %>% 
  filter(pollutant_code == "8001589") %>% 
  group_by(name, address) %>% 
  summarize(mean_pounds = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  ungroup %>% 
  mutate(percent_rank = round(percent_rank(mean_pounds), 2))
```


### Hazardous but not cancerous pollutants 
#### Sulfur dioxide 

```{r}
#| output: true
datatable(cook_county_so2_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```


#### Nitrogen Oxides 

```{r}
#| output: true
datatable(cook_county_nox_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```
#### Phthalic anhydride

```{r}
#| output: true
datatable(cook_county_panhy_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```


#### Maleic anhydride 

```{r}
#| output: true
datatable(cook_county_manhy_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))

  


```


### Known or possible carcinogens 


#### Volatile organic material
- An umbrella category that can include known carcinogens, but as a category is not a carcinogen itself 


```{r}
#| output: true
datatable(cook_county_vom_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```


#### Benzene
- [Well known carcinogen](https://www.epa.gov/sites/default/files/2016-09/documents/benzene.pdf)


```{r}
#| output: true
datatable(cook_county_benzene_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```


#### Naphthalene
- Possible [carcinogen by EPA](https://www.epa.gov/sites/default/files/2016-09/documents/naphthalene.pdf), probable by IARC, known by Cal EPA 


```{r}
#| output: true
datatable(cook_county_naphthalene_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```

#### Quinoline
- [Probable by EPA and IARC](https://iris.epa.gov/ChemicalLanding/&substance_nmbr=1004)


```{r}
#| output: true
datatable(cook_county_quinoline_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```



#### Styrene
- [Probably by IARC](https://www.epa.gov/sites/default/files/2020-05/documents/styrene_update_2a.pdf), but not classified by EPA

```{r}
#| output: true
datatable(cook_county_styrene_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```



#### Creosote
- [Probably by IARC and EPA](https://cfpub.epa.gov/si/si_public_record_Report.cfm?dirEntryID=40306)

```{r}
#| output: true
datatable(cook_county_creosote_avg, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```




## How Koppers compares by select pollutants nationally 

### Data
- National data from the EPA's Toxic Release Inventory on emissions for benzene and naphthalene. This data is the first file of the Basic Plus dataset. I pulled the years 2012 to 2022 since that is what we have for the Cook County emissions inventory. The data loaded here was processed in cleaned in the etl/prep_tri_data.R file. 

### Findings
#### Mean across decade
- Naphthalene: Top 1% of 1,629 emitters 
- Benzene: Top 13% of 1,030 emitters
- Phthalic anhydride: Top 1% (second place) of 127 emitters 
- Maleic anhydride: Top 1% (second place) of 161 emitters
- Quinonline: Top 8% (second place) of 15 emitters
- Styrene: Under bottom 1% of 1,543 polluters 
- Creosote: Around top 60% of 74 polluters 

#### Total across decade
- Naphthalene: Top 1% of 1,629 emitters 
- Benzene: Top 12% of 1,030 emitters
- Phthalic anhydride: Top 1% (second place) of 127 emitters 
- Maleic anhydride: Top 1% (second place) of 161 emitters
- Quinonline: Top 8% (second place) of 15 emitters
- Styrene: Under bottom 1% of 1,543 polluters
- Creosote: Around top 60% of 74 polluters 

```{r}
naphthalene_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "91-20-3") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>%  
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))

benzene_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "71-43-2") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>% 
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))

p_anhydride_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "85-44-9") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>%  
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))

m_anhydride_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "108-31-6") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>% 
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))

quinoline_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "91-22-5") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>%  
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))

styrene_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "100-42-5") %>% 
  group_by(x74_frs_facility_id, x10_facility_name,x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>% 
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))

creosote_sum <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "8001-58-9") %>% 
  group_by(x74_frs_facility_id, x10_facility_name,x11_facility_street, x12_facility_city) %>% 
  summarize(sum_total = sum(x115_total_air_emissions)) %>%  
  filter(sum_total > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(sum_total))


```


```{r}
naphthalene_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "91-20-3") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>% 
  filter(mean > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))

benzene_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "71-43-2") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>%
  filter(mean > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))

p_anhydride_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "85-44-9") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>% 
  filter(mean > 0) %>% 
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))

m_anhydride_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "108-31-6") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>% 
  filter(mean > 0) %>%
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))

quinoline_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "91-22-5") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>% 
  filter(mean > 0) %>%
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))

styrene_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "100-42-5") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>% 
  filter(mean > 0) %>%
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))

creosote_mean <- koppers_tri_pollutants %>% 
  filter(x76_cas_number == "8001-58-9") %>% 
  group_by(x74_frs_facility_id, x10_facility_name, x11_facility_street, x12_facility_city) %>% 
  summarize(mean = round(mean(x115_total_air_emissions),2)) %>% 
  filter(mean > 0) %>%
  ungroup() %>% 
  mutate(percent_rank = percent_rank(mean))


```



#### Naphthalene 

```{r}
#| output: true
datatable(naphthalene_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```


#### Benzene

```{r}
#| output: true
datatable(benzene_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```

#### Phthalic anhydride

```{r}
#| output: true
datatable(p_anhydride_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```

#### Maleic anhydride

```{r}
#| output: true
datatable(m_anhydride_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```


#### Quinoline

```{r}
#| output: true
datatable(quinoline_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```


#### Styrene

```{r}
#| output: true
datatable(styrene_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```



#### Creosote

```{r}
#| output: true
datatable(creosote_mean, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```


#### How do TRI and IEPA data compare 
- A decade of averages line up fairly closely for all pollutants except Benzene, which Koppers reports a lot more of to IEPA than TRI

```{r}
iepa_avgs <- cook_county_emissions %>% 
  filter(name == "Koppers Inc" & address == "3900 S Laramie Ave") %>% 
  group_by(chemical_name, pollutant_code) %>% 
  summarize(iepa_mean = round(mean(emissions, na.rm = TRUE)*2000,2)) %>% 
  mutate(chemical_name = case_when(pollutant_code == "8001589" ~ "Creosote", TRUE ~ chemical_name)) %>% 
  filter(chemical_name %in% c("Naphthalene", "Benzene", "Phthalic anhydride", "Maleic anhydride", "Quinoline", "Styrene", "Creosote"))

tri_avgs <- koppers_tri_pollutants %>% 
  filter(x10_facility_name == "KOPPERS INC", x11_facility_street == "3900 S LARAMIE AVE") %>% 
  group_by(x78_chemical_name) %>% 
  summarize(tri_mean = round(mean(x115_total_air_emissions),2)) %>% 
  rename(chemical_name = x78_chemical_name)

tri_iepa_diff <- iepa_avgs %>% 
  left_join(tri_avgs, by = "chemical_name")

```


```{r}
#| output: true
datatable(tri_iepa_diff, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```





## Cancer Risk by Census tract for benzene and napthalene 
### Data 
- This data comes from the EPA's AirToxScreen and represents the additional amount of people who could have cancer from this specific pollutant, e.g. an extra 10 in one million 
- AirToxScreen includes naphthalene even though it's just a possible carcinogen

### Findings 
- For cancer risk from benzene, the census tracts around Koppers are in top 10% of the country
  - The national average is 1.9 
  - Illinois averages 2 
  - Cicero tracts average 3.04
- For cancer risk from naphthalene, the census tracts around Koppers are in the top 1% of the country. 
  - The national average is 0.8
  - Illinois averages  is 0.7
  - Cicero tracts average 5.69, though the highest is 10, right behind Koppers

### Benzene
```{r}

cicero_tracts <- c("031815500","031814500","031814400", "31814300", "031814200","031820700", "031560100", "031560200", "031820800")

benzene_risk_national <- cancer_risk_tract %>% 
  select(-naphthalene) %>% 
  filter(!county == c("Entire State", "Entire US"))

benzene_risk_cicero <- cancer_risk_tract %>% 
  select(-naphthalene) %>% 
  filter(!county == c("Entire State", "Entire US")) %>% 
  mutate(percent_rank = percent_rank(benzene)) %>% 
  filter(census_tract %in% cicero_tracts)

benzene_risk_illinois <- cancer_risk_tract %>% 
  select(-naphthalene) %>% 
  filter(!county == c("Entire State", "Entire US")) %>% 
  filter(state == "IL") %>% 
  mutate(percent_rank = percent_rank(benzene)) 


national_benzene_risk_avg <- mean(benzene_risk_national$benzene) # is 1.9 for the country
illinois_benzene_risk_avg <- mean(benzene_risk_illinois$benzene) # 2 for illinois 
ciero_benzene_risk_avg <- mean(benzene_risk_cicero$benzene) # is 3.04 for Cicero

```




#### Benzene risk in Cicero tracts near Koppers

```{r}
#| output: true
datatable(benzene_risk_cicero, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```

#### Naphthalene
```{r}

cicero_tracts <- c("031815500","031814500","031814400", "31814300", "031814200","031820700", "031560100", "031560200", "031820800")


naphthalene_risk_national <- cancer_risk_tract %>% 
  select(-benzene) %>% 
  filter(!county == c("Entire State", "Entire US"))

naphthalene_risk_top_100 <- naphthalene_risk_national %>% 
  mutate(percent_rank = percent_rank(naphthalene)) %>% 
  slice_max(order_by = naphthalene, n = 100)


naphthalene_risk_cicero <- cancer_risk_tract %>% 
  select(-benzene) %>% 
  filter(!county == c("Entire State", "Entire US")) %>% 
  mutate(percent_rank = percent_rank(naphthalene)) %>% 
  filter(census_tract %in% cicero_tracts)

naphthalene_risk_illinois <- cancer_risk_tract %>% 
  select(-benzene) %>% 
  filter(!county == c("Entire State", "Entire US")) %>% 
  filter(state == "IL") %>% 
  mutate(percent_rank = percent_rank(naphthalene)) 


national_naphthalene_risk_avg <- mean(naphthalene_risk_national$naphthalene) # is 1.9 for the country
illinois_naphthalene_risk_avg <- mean(naphthalene_risk_illinois$naphthalene) # 2 for illinois 
ciero_naphthalene_risk_avg <- mean(naphthalene_risk_cicero$naphthalene) # is 3.04 for Cicero



```

#### Top 100 Census tracts as for Naphthalene risk across the country

```{r}
#| output: true
datatable(naphthalene_risk_top_100, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```
#### Naphthalene risk in Cicero tracts near Koppers

```{r}
#| output: true
datatable(naphthalene_risk_cicero, filter = 'top', class = 'cell-border stripe order-column', extensions = 'Buttons',

                            options = list(
                                paging = TRUE,
                                searching = TRUE,
                                fixedColumns = TRUE,
                                autoWidth = TRUE,
                                ordering = TRUE,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            ))




```